---
title: "Week 2"
author: "Mathew Merryweather"
date: "16 August 2015"
output: html_document
---

## MySQL
Connecting and Listing databases
```{r}
library(RMySQL)
mysqlURL = "genome-mysql.cse.ucsc.edu"
## dbConnect can connect to multiple DB types
## Always assign connection to object
ucscDB = dbConnect(MySQL(),user = "genome",host = mysqlURL)

## Send a MySQL SQL command
result = dbGetQuery(ucscDB,"show databases;")
dbDisconnect(ucscDB)

## List DB's on server
head(result,5)
```

Connect to hg19 Database and listing tables
```{r}
## 19th build of the human genome
hg19 = dbConnect(MySQL(),user="genome",db="hg19",host=mysqlURL)
allTables = dbListTables(hg19)
## Over 10,000 tables!
length(allTables)
allTables[1:5]
```

```{r}
## One microarray
dbListFields(hg19,"affyU133Plus2")

#return number of records from table
dbGetQuery(hg19,"Select count(*) from affyU133Plus2")
```

Read from the table
```{r}
addyData = dbReadTable(hg19, "affyU133Plus2")
head(addyData)
```

Selecting a specific subset
```{r}
## dbSendQuery Allows you to subset data
query = dbSendQuery(hg19,"select * from affyU133Plus2 where misMatches between 1 and 3")
affyMis  = fetch(query)
quantile(affyMis$misMatches)

## Suck out a small amount using fetch
affyMisSmall = fetch(query,n=10)
## Must clear 
dbClearResult(query)
dim(affyMisSmall)
dbDisconnect(hg19)
```

Never, never, NEVER perform

* deletes
* adds
* joins

to public facing DBs. Only select, ever.

***

### HDF5
Heirachial data format
For storing large datasets
Groups containing zero or mor

R HDF5
```{r}
source("http://bioconductor.org/biocLite.R")
biocLite("rhdf5")
# Genomics and big data tools
```

```{r}
library("rhdf5")
created = h5createFile("example.h5")
created
```

create groups
```{r}
eg = "example.h5"
created = h5createGroup(eg,"foo")
created = h5createGroup(eg,"baa")
created = h5createGroup(eg,"foo/foobaa")
h5ls(eg)
```

write to groups
```{r}
a = matrix(1:10,nr=5,nc=2)
h5write(a,eg,"foo/A")
b = array(seq(0.1,2.0,by=0.1),dim=c(5,2,2))
attr(b,"scale") = "liter"
h5write(b,eg,"foo/foobaa/b")
h5ls(eg)
```

read data
```{r}
readA = h5read(eg, "foo/A")
readB = h5read(eg, "foo/foobaa")
```

writing/reading chunks
```{r}
## Index can be used for reading as well
h5write(c(12,13,14),eg,"foo/A",index=list(1:3,1))
h5read(eg,"foo/A")
```








