---
title: "Week-1-Lectures"
author: "Mathew Merryweather"
date: "8 September 2015"
output: html_document
---

## Principles of analytic graphs
general rules (Tuftian)

Principle 1

* Show comparisons (usually basline or control)
* evidence is relative to another competeing hypothesis
* "Compared to what?"

Principle 2

* show causality, mechanism, explanation or systematic structure
* what is your causual framework for thinking about a question?
* how is the system operating?
* Graph 1 = Outcome, Graph 2 = Change in input

Principle 3

* Show multivariate data
* Real world is multivariate
* Need to escape flatland
* Multivariate analysis can have a very large effect (Symptoms effect)

Principle 4

* Integration of evidence
* Completely integrate words, numbers etc
* Make use of many modes
* Dont let the tools drive the analysis

Principle 5

* Describe and document the evidence with labels, scales and source
* Include code

Principal 6

* Content is king!
* Quality, relevance and integrity of their content.

## Exploratory Graphs

More or less for yourself

* To understand data properties
* finding patterns
* Debug analysis
* proposing modelling strategies

* Made quickly
* large number,
* goal is personal understanding
* Colour / size used for info

## Simple summaries of Data: 1D

* Six-number summary `summary`
* Boxplots `boxplot`
    * `abline` to include a line (`h` or `v` for the direction)
* Histograms `hist` + `rug` + `abline`
    * Look at `breaks` argument
* Density Plot 
* Barplot `barplot` for categorical variables
 
## Simple summaries of Data: 2D

2 dimensions

* Multiple/overlayed 1-D plots (`lattice`/`ggplot2`)
* Scatter plots
* Smooth Scatter plots

>2 dimensions
* overlayed 2-D plots: `coplots`
* user colour / shape
* Spinning plots

multiple histograms
```{r eval = F}
par(mfrow = c(2,1), mar = c(4,4,2,1))
hist(subset(pollution,region=="east")$pm25)
hist(subset(pollution,region=="west")$pm25)
```

scatterplot
```{r eval = F}
with(pollution,plot(latitude,pm25))
abline(h=12,lwd=2,lty=2)
```

scatter with colour
```{r eval = F}
with(pollution,plot(latitude,pm25,col=region))
abline(h=12,lwd=2,lty=2)
```

multiple scatterplots
```{r eval = F}
par(mfrow = c(1,2), mar = c(5,4,2,1))
with(subset(pollution,region=="West"),plot(latitude, pm25, main = "West"))
with(subset(pollution,region=="east"),plot(latitude, pm25, main = "East"))
```


