---
title: "Week-1-Lectures"
author: "Mathew Merryweather"
date: "8 September 2015"
output: html_document
---

## Principles of analytic graphs
general rules (Tuftian)

Principle 1

* Show comparisons (usually basline or control)
* evidence is relative to another competeing hypothesis
* "Compared to what?"

Principle 2

* show causality, mechanism, explanation or systematic structure
* what is your causual framework for thinking about a question?
* how is the system operating?
* Graph 1 = Outcome, Graph 2 = Change in input

Principle 3

* Show multivariate data
* Real world is multivariate
* Need to escape flatland
* Multivariate analysis can have a very large effect (Symptoms effect)

Principle 4

* Integration of evidence
* Completely integrate words, numbers etc
* Make use of many modes
* Dont let the tools drive the analysis

Principle 5

* Describe and document the evidence with labels, scales and source
* Include code

Principal 6

* Content is king!
* Quality, relevance and integrity of their content.

## Exploratory Graphs

More or less for yourself

* To understand data properties
* finding patterns
* Debug analysis
* proposing modelling strategies

* Made quickly
* large number,
* goal is personal understanding
* Colour / size used for info

## Simple summaries of Data: 1D

* Six-number summary `summary`
* Boxplots `boxplot`
    * `abline` to include a line (`h` or `v` for the direction)
* Histograms `hist` + `rug` + `abline`
    * Look at `breaks` argument
* Density Plot 
* Barplot `barplot` for categorical variables
 
## Simple summaries of Data: 2D

2 dimensions

* Multiple/overlayed 1-D plots (`lattice`/`ggplot2`)
* Scatter plots
* Smooth Scatter plots

>2 dimensions
* overlayed 2-D plots: `coplots`
* user colour / shape
* Spinning plots

multiple histograms
```{r eval = F}
par(mfrow = c(2,1), mar = c(4,4,2,1))
hist(subset(pollution,region=="east")$pm25)
hist(subset(pollution,region=="west")$pm25)
```

scatterplot
```{r eval = F}
with(pollution,plot(latitude,pm25))
abline(h=12,lwd=2,lty=2)
```

scatter with colour
```{r eval = F}
with(pollution,plot(latitude,pm25,col=region))
abline(h=12,lwd=2,lty=2)
```

multiple scatterplots
```{r eval = F}
par(mfrow = c(1,2), mar = c(5,4,2,1))
with(subset(pollution,region=="West"),plot(latitude, pm25, main = "West"))
with(subset(pollution,region=="east"),plot(latitude, pm25, main = "East"))
```

## Base Plotting System in R

Core is contained in two packages

* `graphics`: `plot`,`hist`,`boxplot`
* `grDevices`: X11, PDF, PNG

Lattice is implemented

* `lattice`: trellis graphics, `xyplot`, `bwplot`, `levelplot`
* `grid`: Lattice is built off `grid`. rarely called directly

**Process of making a plot**

Questions

* the audience?
    * quality
    * print or web?
* amount of data?
* Dynamic resize

System to use? Cannot be mixed
* `base`: piece meal, assembled
* `lattice`: single function call
* `ggplot2`: mix of both

Focus on base for screens

**Base**

1. initialize a new plot
2. Annotating existing plot

calling `plot(x,y)` etc will launch graphics devices if ones not already open.
depends on data passed to it.
Default type has lots of params

```{r}
library(datasets)
hist(airquality$Ozone)
```

scatterplot

```{r}
library(datasets)
with(airquality, plot(Wind, Ozone))
```

Boxplots

```{r}
airquality = transform(airquality, Month = factor(Month))
boxplot(Ozone ~ Month, airquality, xlab = "Month", ylab = "Ozone (ppb)")
```

Key parameters for base

* `pch`: Plotting Character
* `lty`: Line Type
* `lwd`: Line Width
* `col`: color (Contrast with `colors()`)
* `xlab`
* `ylab`

global parameters using `par`
* `las`: label orientation
* `bg`: background col
* `mar`: margin size
* `oma`: outer margin size
* `mfrow`: number of plots per row, col 

## Base Plotting Functions

* `plot`: Scatter plot
* `lines`: connects dots of scatter plot
* `points`: add points to a plot
* `text`: add text labels to a plot
* `title`: adds annotations
* `mtext`: adds arbitary text to margins
* `axis`: axis ticks / labels

```{r}
with(airquality,plot(Wind,Ozone,main = "Ozone and Wind in NYC"))
with(subset(airquality, Month == 5), points(Wind,Ozone, col = "blue"))
```

```{r eval = F}
plot(type = "n")
## Doesn't draw graph
legend("topright",pch=1,col=c("blue","red"),legend = c("May","Other"))
```

Base Plot with Regression Line
```{r}
with(airquality, plot(Wind,Ozone, main = "Ozone and Wind in NYC", pch = 20))
model = lm(Ozone ~ Wind, airquality)
abline(model, lwd = 2)
```

Multiple Base Plots
```{r}
par(mfrow=c(1,2), oma = c(0,0,2,0))
with(airquality,{
    plot(Wind, Ozone, main = "Ozone and Wind")
    plot(Solar.R, Ozone, main = "Ozone and Solar Radiation")
    mtext("Ozone and Weather in NYC", outer = TRUE)
})
```

