---
title: "Week 3 Lectures"
author: "Mathew Merryweather"
date: "17 August 2015"
output: html_document
---

## Subsetting

```{r}
set.seed(13435)
x = data.frame("var1"=sample(1:5),"var2"=sample(6:10,"var3"=sample(11:15))) 
```


## Summarising
```{r}
if(!file.exists("./data")){dir.create("./data")}
fileUrl = "https://data.baltimorecity.gov/api/views/k5ry-ef3g/rows.csv?accessType=DOWNLOAD"
download.file(fileUrl,destfile = "./data/restaurants.csv",method="curl")
data = read.csv(file = "./data/restaurants.csv")
```

summary functions
```{r}
summary(data)
str(data)

#Quantiles of quantitive variables
quantile(data$councilDistrict,na.rm=T)
quantile(data$councilDistrict,probs=c(0.5,0.75,0.9))
```

Make tables
```{r}
#deafault is to hide
table(data$zipCode,useNA = "ifany")
```

```{r}
# two way
table(data$councilDistrict, data$zipCode)
```

Missing values
```{r}
sum(is.na(data$councilDistrict))
any(is.na(data$councilDistrict))
all(data$zipCode>0)

# By row or column
colSums(is.na(data))
all(colSums(is.na(data))==0)
```

values with specific characteristics
```{r}
table(data$zipCode %in% c("21212"))
table(data$zipCode %in% c("21212","21213"))

## Creates logical vector
head(data[data$zipCode %in% c("21212","21213"),],5)
```

Cross Tabs
```{r}
data("UCBAdmissions")
DF = as.data.frame(UCBAdmissions)
summary(DF)
xt = xtabs(Freq ~ Gender + Admit, data=DF)
xt
```

Flat Tables
```{r}
warpbreaks$replicate = rep(1:9,len = 54)

# Breakdown by everything else
xt = xtabs(breaks ~.,data=warpbreaks)
xt
ftable(xt)
```

Size of a data set
```{r}
fakeData = rnorm(1e5)
object.size(fakeData)
print(object.size(fakeData),units="Mb")
```

***

## Reshaping Data
The goal is tidy data

1. Each varable forms a column
2. Each observation forms a row
3. Each table/file stores data about one kind of observation (e.g people, hospitals, states)

start with reshaping
```{r}
library(reshape2)
head(mtcars,5)
## Assign rownames to have an explicit column
mtcars$carname = rownames(mtcars)
carMelt = melt(mtcars, id = c("carname","gear","cyl"),measure.vars = c("mpg","hp"))
head(carMelt,3)
```
Casting data frames
```{r message=FALSE}
## Puts cylinders in rows, other vars in cols
## Will default to length / count
cylData = dcast(carMelt, cyl ~ variable)
cylData
```

Averaging values
```{r}
## Summarise by mean, similar to Pivot Table
cylData = dcast(carMelt, cyl ~ variable, mean)
cylData
```

```{r}
head(InsectSprays)
## Sum of the counts of each spray
## Apply along the index of Sparays, sum the count
tapply(InsectSprays$count,InsectSprays$spray,sum)
```

Using the split-apply-combine method
```{r}
## SPLIT into a list
spIns = split(InsectSprays$count,InsectSprays$spray)
spIns

## APPLY sums for each element of the list
sprCount = lapply(spIns, sum)
sprCount

## COMBINE the results into a result 
unlist(sprCount)
## OR in fewer steps (Apply + Combine)
sapply(spIns, sum)
```

Using plyr
```{r}
library(plyr)
## SPLIT APPLY COMBINE in one function
## Take insect sprays, summarize for each .(spray)
## Summarize this var
## How? by sum of the counts
ddply(InsectSprays,.(spray),summarize,sum=sum(count))
```

Nice way of creating summary variables
```{r}
## Creating a variable that has the means for each spray in each variable.
spraySums = ddply(InsectSprays,.(spray),summarize,sum=ave(count,FUN = sum))
dim(spraySums)
## By passing ave with the a subfunction of the sum
## See the sum of A for each variable
```

## Managing data with dplyr

* arrange
* filter
* select
* mutate
* rename

dplyr is the optimized and distilled dplyr
Just simpler and faster

dplyr verbs

* `select:` return a subset of the cols in a df
* `filter:` extract a subset of rows from a df based on logical conditions
* `arrange:` reorder rows of a df
* `rename:` rename variables in a df
* `mutate:` add new vars/cols or transform existing
* `summarise:` generates summary statistics in the data frame

Properties:
1. First argument is *always* a data frame
2. the subsequent args describe what to do with it. Don't have to use the $ operator, just names
3. The result is *always* a new data frame

Hence the importance of usefully named and annotated data frames


## Working with dplyr
`%>%` is the pipeline operator, implicitly passes resulting dataframe to next function. Allows you to chain together functions without temporary variables

`dplyr` grammar can also be used with `data.table` or `SQL` as backends.


## Merging Data
Matching based on ID















