---
title: "Week 1 Quiz"
author: "Mathew Merryweather"
date: "10 August 2015"
output: html_document
---

#### Q1
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv 

and load the data into R. The code book, describing the variable names is here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf 

How many properties are worth $1,000,000 or more?
```{r}
filePath = "Q1.csv"
if(!file.exists(filePath))
{
    fileURL = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
    download.file(fileURL,filePath, method = "curl")   
}
df = read.csv(filePath)
val = df$VAL
Q1 = sum(val==24, na.rm = T)
Q1
```

#### Q2
Use the data you loaded from Question 1. Consider the variable FES in the code book. Which of the "tidy data" principles does this variable violate?
```{r}
filePath = "Q1.csv"
if(!file.exists(filePath))
{
    fileURL = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
    download.file(fileURL,filePath, method = "curl")   
}
df = read.csv(filePath)
val = df$FES
str(val)
#Tidy data has one variable per column.
```

#### Q3
Download the Excel spreadsheet on Natural Gas Aquisition Program here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx 

Read rows 18-23 and columns 7-15 into R and assign the result to a variable called:
```{r eval=F}
dat
```
What is the value of:
```{r eval=F}
sum(dat$Zip*dat$Ext,na.rm=T) 
```
```{r eval = T}
filePath = "Q3.xlsx"
if(!file.exists(filePath))
{
    fileURL = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
    download.file(fileURL,filePath, method = "curl")   
}

rowIndex = 18:23
colIndex = 7:15
library(xlsx)
dat = read.xlsx(filePath, rowIndex = rowIndex, colIndex = colIndex, sheetIndex = 1)
val =  sum(dat$Zip*dat$Ext,na.rm=T) 
val
```

#### Q4
Read the XML data on Baltimore restaurants from here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml 

How many restaurants have zipcode 21231?

```{r}
library(XML)

filePath = "Q4.xml"
if(!file.exists(filePath))
{
    fileURL = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"
    download.file(fileURL,filePath, method = "curl")   
}
doc = xmlTreeParse(filePath, useInternal = T)
rootNode = xmlRoot(doc)
df = xpathSApply(rootNode,"//zipcode",xmlValue)
zipTarget = 21231
Q4 = sum(df == zipTarget, na.rm = T)
Q4
```

#### Q5
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv 

using the fread() command load the data into an R object
```{r eval = F}
 DT 
```

Which of the following is the fastest way to calculate the average value of the variable
```{r eval = F}
pwgtp15 
```

broken down by sex using the data.table package?

```{r}
library(data.table)
filePath = "Q5.csv"
if(!file.exists(filePath))
{
    fileURL = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
    download.file(fileURL,filePath, method = "curl")   
}
DT = fread(filePath)
x1 = system.time(mean(DT$pwgtp15,by=DT$SEX))
x2 = system.time(DT[,mean(pwgtp15),by=SEX])
x3 = system.time(tapply(DT$pwgtp15,DT$SEX,mean))
x4 = system.time(mean(DT[DT$SEX==1,]$pwgtp15)); mean(DT[DT$SEX==2,]$pwgtp15)
x5 = system.time(sapply(split(DT$pwgtp15,DT$SEX),mean))
times = rbind(x1,x2,x3,x4,x5)
times
```

 


