---
title: "Week 3 Quiz"
author: "Mathew Merryweather"
date: "28 July 2015"
output: html_document
---

## Q1
In this dataset, what is the mean of 'Sepal.Length' for the species *virginica*?
```{r}
library(datasets)
data(iris)
virginica = split(iris,iris$Species)$virginica
Q1 = mean(virginica$Sepal.Length)
Q1
```

## Q2
Continuing with the 'iris' dataset from the previous Question, what R code returns a vector of the means of the variables 'Sepal.Length', 'Sepal.Width', 'Petal.Length', and 'Petal.Width'?
```{r}
apply(iris[,1:4],2,mean)
```

## Q3
How can one calculate the average miles per gallon (mpg) by number of cylinders in the car (cyl)?
```{r}
#rm(list = ls())
library(datasets)
data(mtcars)
sapply(split(mtcars$mpg, mtcars$cyl), mean)
```

## Q4
Continuing with the 'mtcars' dataset from the previous Question, what is the absolute difference between the average horsepower of 4-cylinder cars and the average horsepower of 8-cylinder cars?
```{r}
hp = sapply(split(mtcars$hp,mtcars$cyl), mean)
Q4 = as.numeric(hp[3]-hp[1])
Q4
```

## Q5
If you run
```{r}
debug(ls)
```
what happens when you next call the 'ls' function?
```{r}
print("Execution of 'ls' will suspend at the beginning of the function and you will be in the browser.")
```



