---
title: "Week3 Lecture"
author: "Mathew Merryweather"
date: "27 July 2015"
output: html_document
---

##Lapply
Looping on the CLI

* `lapply:` Loop over a list and evaluate a function on each element
* `sapply:` Same as `lapply` but try to simplify results
* `apply:` Apply a function of the margins of an array (useful for summaries)
* `tapply:` Apply a function over subsets of a vector (table apply)
* `mapply:` Multivariate version of lapply

Auxillarilary function is `split`, used often in conjection with `lapply`. Splits objects into subpieces

#lapply
Three args:
* A list `x`
* A function `FUN`
* Other arguments via its arguments

`x` will be coerced to a list using as.list

```{r}
x = list(a = 1:5, b = rnorm(10))
lapply(x, mean)
#NB: Same names
x$a
x$b
```

```{r}
x = 1:4
lapply(x, runif)
```
Passing arguments
```{r}
x  = 1:4
lapply(x, runif, min = 0, max = 10)
```

makes heavy use of *anonymous* functions

```{r}
x = list(a = matrix(1:4,2,2),b = matrix(1:6,3,2))
x
#No function to extract first col
# so we will write one. elt is not found anywhere else
lapply(x,function(elt) elt[,1])
```

##sapply
lapply always returns list.
If the result is a list where every element is length 1 then a vector is returned.
If the result is a list where every elemtn is a vector of the same length (>1) a matrix is returned

```{r}
x = list(a = 1:4, b = rnorm(10), c= rnorm(20,1), d = rnorm(100,5))
sapply(x,mean)
```

##apply
often applied to rows or columns. Can take a general array
```{r}
str(apply)
```
notes:

*`x` is an array
* `MARGIN` is an integer vector indicating which margins should be retained
* `FUN` is a function to be applied
* `...` as before, to pass args


```{r}
x = matrix(rnorm(200),20,10)
#Calculate for x, each columns mean
# second dim is the cols (rows > cols)
# Collapse the first dimension
apply(x,2,mean)
apply(x,1,sum)
```

Shortcuts: Much faster!

* `rowSums` = `apply(x,1,sum)
* `rowMeans` = apply(x,1,mean)
* `colSums` = apply(x,2,sum)
* `colMeans` = apply(x,2,mean)

```{r}
#Lets say you have a matrix
# and you need to calculate the
# 25th and 75th percentile for each row

x = matrix(rnorm(200),20,10)
#Remember, keep 1st rows > cols
apply(x,1,quantile, probs = c(0.25, 0.75))
```

condensing higher dimensions
```{r}
x = array(rnorm(2 * 2 * 10), c(2,2,10))
apply(x, c(1,2), mean)
rowMeans(x,dims=2)
```

##Mapply
If you needed elements from multiply lists (of equal length), instead of a for loop with two indexes, use `mapply`
```{r}
#first arg from first list
mapply(rep,1:4,4:1)
```

```{r}
noise = function(n,mean,sd){
    rnorm(n,mean,sd)
}
noise(5,1,2)
#Changing n and mean
mapply(noise,1:5,1:5,2)
```
Good for instant vectorization

##tapply
Useful for apply to only a subset of items
```{r}
str(tapply)
```

Notes:

* `x:` is a vector
* `INDEX:` is a factor or a list of factors
* `FUN:` function or summary statistic to be applied
* `...` arguments to be passed to `FUN`
* `simplify:` try and simplify the result to a vector or matrix

```{r}
x = c(rnorm(10),runif(10),rnorm(10,1))
f = gl(3,10)
f
tapply(x, f, mean, simplify = FALSE)
tapply(x, f, range)
```

***

##split
split takes a vector or other objects and splits it into groups determined by a factor or list of factors

* `x` is a vector/list/dataframe
* `f` is a factor
* `drop` indicates whether empty factor levels should be dropped

```{r}
x = c(rnorm(10),runif(10),rnorm(10,1))
f = gl(3,10)
split(x,f)
```

A common pattern is `split` followed by `lapply`
```{r}
lapply(split(x,f),mean)
#THATS FUCKING COOL
```
Analagous to `tapply` but `split` is much more useful for complex objects


```{r}
library(datasets)
head(airquality)

#summary for each month
s = split(airquality,airquality$Month)
# Anonymous function for colmeans of only 3 cols.
lapply(s, function(x) colMeans(x[,c("Ozone","Solar.R","Wind")]))
```

Using `sapply` becuase `lapply` returns a list
```{r}
sapply(s, function(x) colMeans(x[,c("Ozone","Solar.R","Wind")]))
sapply(s, function(x) colMeans(x[,c("Ozone","Solar.R","Wind")],na.rm = T))
```

splitting on more than one level
```{r}
x = rnorm(10)
f1 = gl(2,5)
f2 = gl(5,2)
#Combines levels of first and second
interaction(f1,f2)
```

```{r}
#Calling interaction implicitly
str(split(x,list(f1,f2)))
#Drop empty levels created from splitting
str(split(x,list(f1,f2),drop = TRUE))
```















