---
title: "Week 4 Lectures"
author: "Mathew Merryweather"
date: "30 July 2015"
output: html_document
---

#### The Str Function
A diagnostic function, alternative to `summary`
Compactly display lists. One line per basic object.
"Whats in this object?"
```{r}
str(str)
str(lm)
```

```{r}
x = rnorm(100,2,4)
summary(x)
str(x)

#Factor variable
f = gl(40,10)
str(f)
```

```{r}
library(datasets)
head(airquality)
str(airquality)
```

```{r}
m  = matrix(rnorm(100),10,10)
str(m)
```

```{r}
s = split(airquality, airquality$Month)
str(s)
```

***

## Simulation

#### Generating Random Numbers
functions for probability distributions in R

* `rnorm`: Generate random Normal variates with a given mean and standard of deviation
* `dnorm`: evaluate the normal probability density (with a given mean/SD) at a point (or a vector of points)
* `pnorm`: evaluate the cumulative distribution function for a normal distribution
* `rpois`: Generate random Poisson variates with a given rate (integers)

PDFs usually have four functions associated with them

* `d` for density
* `r` for random number generation
* `p` for cumulative distribution
* `q` for quantile functions

i.e. `dgamma`, `pgamma` etc.

All require mean and SD, if none provided then Standard Distribution
```{r eval = FALSE}
#Often need log of density
dnorm(x, mean = 0, sd = 1, log = FALSE)
#Also log scale option
#Lower tail = Left, 
#Lower.Tail = False = right side
pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)
qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)
rnorm(n, mean = 0, sd = 1)
```

if $\Phi$ is the CDF for a standard normal distribution, then `pnorm(q)` = $\Phi$`(q)` and `qnorm(p)` = $\Phi^-1$ `(p)`.

```{r}
#ALWAY SET SEED!!!
set.seed(1)
x = rnorm(10)
x
x = rnorm(10, 20,2)
x
summary(x)
```

Always use the `set.seed` function to ensure reproducibility
```{r}
set.seed(1)
rnorm(5)
rnorm(5)
set.seed(1)
rnorm(5)
```

Generating Poisson Data

```{r}
rpois(10,1)
rpois(10,2)
rpois(10,20)

# Mean = Rate

# Probability that a poisson value is less than 2 when I have a rate of 2?
# Use Cumulative distribution
ppois(2,2) ## Pr(x <= 2)
ppois(4,2) ## Pr(x <= 4)
ppois(6,2) ## Pr(x <= 6)
```

#### Simulating a linear model
Suppose we want to simulate from the following Linear Model
$$
y = \beta_{0} + \beta_{1}x + \epsilon
$$

Where $\epsilon \sim \mathcal{N} (0,2^2)$ Assume $x \sim \mathcal{N} (0,1^{2})$, $\beta_{0} = 0.5$ and $\beta_{1} = 2$

```{r}
# Important
set.seed(20)
# Predictor
x = rnorm(100)
# Errpr
e = rnorm(100,0,2)
# Output
y = 0.5 + 2 * x + e
summary(y)
plot(x,y)
```

what if `x` is a binary?
perhaps gender or taken medication vs not?

```{r}
set.seed(10)
# 100 variables, size and probability
# i.e Male vs Female
x= rbinom(100,1,0.5)
e = rnorm(100,0,2)
y = 0.5 + 2 * x + e
summary(y)
plot(x,y)
```

Suppose we want to simulate from a Poisson model where

$Y \sim Poisson(\mu)$

$log(\mu) = \beta_{0} + \beta_{1}x$

and $\beta_{0} = 0.5$ and $\beta_{1} = 0.3$

```{r}
set.seed(1)
# Outcome variables for count data
# Error distribution is now a Poisson

#Predictor
x= rnorm(100)
#Linear Predictor
log.mu = 0.5 + 0.3 * x
#To get mean, we need exponent of log.mu
y = rpois(100,exp(log.mu))

summary(y)
plot(x,y)
```

***

#### Random Sampling
`sample` function draws randomly from vector of numbers

```{r}
set.seed(1)
#No replacement
sample(1:10,4)
sample(1:10,4)

sample(letters,5)
sample(1:10) ## Permutation

sample(1:10, replace = T) ## Smaple with replacement
```

***

#### Profiler
Useful for identifying bottleknecks and identifying slow points
Other tools also help to optimize software.
Don't think about it at first. (Knuths Truths)
Design first then optimize
Collect data, don't guess!

Using system.time()
Returns time to evaluate expression in seconds
two types of time:

* *User Time* = CPU Time
* *Elapsed Time* = "wall clock time"

Elapsed time can be greater if CPU spends a lot of time waiting around.
Elapsed time can be shorter if you have multiple cores.

Using `Rprof()`
The summaryRprof() function summarizes the output from `Rprof`

DO NOT USE `system.time()` and `Rprof` together.

Queries the call stack at a sampling interval (default 0.02 s)

Two methods of harnessing `summaryRprof`

* `by.total` total run time 
* `by.self` subtracts out time below in the call stack.







