---
title: "Week 2 Optimisations"
author: "Mathew Merryweather"
date: "15 July 2015"
output: html_document
---

### Application Optimization
`optim`,`nlm`,`optimize` require that you pass a function whose argument is a vector of parameters.

i.e. you pass these functions a function.

But if it needs data as well, how do you do it?

You write a constructr function, constructs objective function. All the data is in the objective functions environment.

```{r eval = T}
## NOTE:
## Optimization  functions in R *minimize* functions
## We need to use the negative log-likelihood to find the max of the log-likelihood

make.negLogLik = function(data, fixed=c(F,F)){
  params = fixed
  function(p){
    params[!fixed] = p
    mu = params[1]
    sigma = params[2]
    
    a = -0.5 * length(data)*log(2*pi*sigma^2)
    b = -0.5 * sum((data - mu)^2)/ (sigma^2)
    -(a+b)
  }
}
```

The fixed parameter tells us if we want to change or keep parameters
```{r eval = T}
set.seed(1); normals = rnorm(100,1,2) # mean 1, SD = 2
#nLL = Negative Log Likelihood
nLL = make.negLogLik(normals)
nLL
#Note Hexadecimal address
ls(environment(nLL))
```

By using the consturctor function, the "data" variable is available, it doesn't need to be specified each time.

```{r}
optim(c(mu =0, sigma =1),nLL)$par
```

Close to originals of 1 and 2

fixing $\sigma$ = 2
```{r}
#Single variable so can use optimize instead of optim
nLL = make.negLogLik(normals, c(F,2))
optimize(nLL,c(-1,3))$minimum
```

fixing $\mu$ = 1
```{r}
nLL = make.negLogLik(normals, c(1, F))
optimize(nLL, c(1e-6,10))$minimum
```

Plotting the likelihood
```{r eval = T}
nLL = make.negLogLik(normals, c(1,F))
x = seq(1.7,1.9, len = 100)
y = sapply(x, nLL)
plot(x, exp(-(y-min(y))), type = "l")

```



