---
title: "Week 2"
author: "Mathew Merryweather"
date: "14 July 2015"
output: html_document
---
#Control Structures
Similar to many other languages. Not recommended for interactive mode. Better to use the *apply functions instead.

* if / else
* for
* while
* break (exit a loop)
* next (skip iteration of loop)
* return (exit a function)

##If/Else
Structure is:
```{r eval=F}
if(condition){
    ## Foo
}else if(condition2){
    ## Bar
}else{
    ## Boo
}
```
Can assign variables in the loop or as the argument to variable assignment

##For Loops
Structure is:
```{r eval = T}
x = c("a","b","c","d")

#Classic Method
for(i in 1:4){
    print(x[i])
}

#seq_along Method (Creates integer sequence)
for(i in seq_along(x)){
    print(x[i])
}

#Compact Method for single method
for(i in 1:4) print(x[i])
```

Useful functions:

* `seq_len`
* `seq_along`
* `nrow` and `ncol`

## While Loops
```{r eval=T}
#Single condition
count =0
while(count<3){
    print(count)
    count = count + 1
}

#Multiple condition
z = 5
while(z >= 3 && z <=10){
    print(z)
    coin = rbinom(1,1,0.5)
    if(coin ==1){ #Random Walk
        z = z +1
    }else{
        z = z-1
    }
}

```

## Functions
Of the form:
```{r eval = F}
foobar = function(a, b = NULL){
    #b = Null assigns a default value to the variable b
    a^2
    #R automatically returns the last line of code
}

foobar(2)
```

Arguments are evaluated lazily. Only evaluated as need. foobar above is *lazy*. foobar won't error even though no `b` argument is passed.

Special argument `...`. allows for extending another function and you don't want to copy the entire argument list.
Can also be used for generic functions but can pass methods to objects.
Also used if the number of arguments to be passed is unknown in advance.

```{r eval  = T}
args(paste)
# No idea of how many objects you want to paste together

args(cat)
#Concatencate

#Arguments after the `...` MUST be be named, cannot be partially named. Must be explicit
```

## Symbol Binding

```{r}
lm = function(x){x * x
}
lm
```
How does it know to use the local function instead of the ones in the `stats` package?

R tries to search from the highest order environment then works down
```{r}
search()
```

R however has separate namespaces for objects and functions.
Packages have no fixed order, so never rely on it!

### Scoping Rules
R uses lexical scoping over dynamic scoping
```{r}
f = function(x,y){
    x^2 + y / z
    #z is a free variable!
}
```

The values of free variables are searched for in the environment in which the function was defined.
Every environment has a parent environment except for the empty env. Collections of symbols and values.

R lets you have functions defined *inside other functions*.
A function can return *another* function as the result which has a different environment.

```{r}
make.power = function(n){
    #Constructor
    pow = function(x){
        x^n
    }
    pow
}

cube = make.power(3)
square = make.power(2)

cube(3)

```

The `ls` function is used to explore a functions environment

## Dates and Times
Dates are represented by the `Date` class
Time is represented by `POSIXct` or `POSIXlt`

Dates have no times. i.e. days since 1970-01-01
Times are seconds since then.

```{r}
x = as.Date("1970-01-01")
x
unclass(x)
unclass(as.Date("1970-01-02"))
```

## Times in R
Times have two types.
POSIX is family of standards.

The `POSIXct` class is essentially an integer representing seconds since the UNIX epoch whilst the `POSIXlt` class is a list.
the lt class stores day, week, day of year, month etc.

you can coerced between them with the following:

* `as.POSIXlt(x)`
* `as.POSIXct(x)`

POSIXlt has nice access to date parts.

strptime converts strings to times using a format string,

```{r}
datestring = c("January 10, 2012 10:40")
x = strptime(datestring, "%B %d, %Y %H:%M")
## Just use ?strptime to figure out the format
class(x)
```

Make sure you use the same class when trying to perform operations. Date classes handle a lot of the complexities with dates for you.









